Learning Distributed Word Representations for Natural Logic Reasoning

### 2015-01-07 writing/AAAI SSS '15/BowmanPottsManning.tex
AAAI paper updates

Natural logic aligns well with distributed (e.g., vector)
representations, which also naturally model meaning relationally.
Distributed representations have been used successfully in a wide
array of sophisticated language tasks (e.g., \cite{collobert2011natural}). %,
% including sentiment analysis, analogy completion, relation
% extraction, and named entity recognition.
However, it remains an open question whether it is possible to train
such representations to support the rich, diverse logical reasoning
captured by natural logic; while they excel at synonymy (similarity),
the results are more mixed for entailment, contradiction, and mutual
consistency.  Using the natural logic of \cite{maccartney2009extended}
as our formal model, we address this open question using two neural
network-based models for learning embeddings: plain neural networks
and neural tensor networks (NTNs).  The natural logic is built from
the seven relations defined in Table~\ref{b-table}. Its formal
properties are now well-understood \cite{Icard:Moss:2013:LILT}, so it
provides a rigorous set of goals for our neural models. To keep the
discussion manageable, we limit attention to experiments involving the
lexicon; for a more extended treatment of complex expressions
involving logical connectives and quantifiers, see
\cite{Bowman:Potts:Manning:2014}. % CITET

@@

% TODO: Better Gref citation; cite Oriol/Ilya

% In our first experiment, we use the pre-specified logical grammar to
% generate controlled data sets and assess the ability of the two
% classes of neural network to learn the core underlying relational
% algebra from this simulated data. In our second experiment, we define
% a simple modification of the logical grammar so that it aligns with
% the WordNet noun graph, and we again assess our two classes of RNN on
% this data. The controlled data of our first experiment helps us to
% achieve analytic insights into what the models are able to learn, and
% the large, diverse lexicon of the second experiment shows us how they
% fare in the real world.

% These experiments differentiate the increased power of RNTNs better
% than previous work and provide the most convincing demonstration to
% date of the ability of neural networks to model semantic inferences
% in complex natural language sentences.

% TODO: Add citations on related work in structuring relations in embedding spaces:
% - Representing part–whole relations in conceptual spaces
% - Something McCallum/Universal Schema?

\begin{table*}[tp]
  \centering\small
  \setlength{\tabcolsep}{15pt}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabular}{l c l l} 
    \toprule
    Name & Symbol & Set-theoretic definition & Example \\ 
    \midrule
    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle $\natfor$ reptile}  \\ 
    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile $\natrev$ turtle}  \\ 
    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch $\nateq$ sofa} \\ 
    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle $\natalt$ warthog} \\ 
    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able $\natneg$ unable} \\
    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal $\natcov$ non-turtle} \\ 
    independence       & $x \natind y$   & (else) & \ii{turtle $\natind$ pet}\\
    \bottomrule
  \end{tabular}
  \protect\caption{\protect\label{b-table}The seven natural logic relations of \protect\cite{maccartney2009extended}. 
    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
    %is insufficient knowledge to choose a label.}
\end{table*}

@@

We build embedding-based models using the method of
\cite{Bowman:Potts:Manning:2014}, which is centered on the task of
labeling a pair of words or sentences with one of a small set of
logical relations. The architecture that we use, which is limited to
only pairs of single terms (such as words), is depicted in
Figure~\ref{sample-figure}. The model represents the two input terms
as embeddings, which are fed into a comparison function based on one
of two types of neural network layer functions to produce a
representation for the relationship between the two terms. This
representation is then fed into a softmax classifier, which outputs a
distribution over possible labels. The entire network, including the
embeddings, is trained using backpropagation with AdaGrad \cite{duchi2011adaptive}.

The simpler version of the comparison concatenates the two
input vectors before feeding them into a standard neural network (NN)
layer.  The more powerful neural tensor network (NTN) version uses an
additional third-order tensor parameter to allow for multiplicative
interactions between the two inputs \cite{chen2013learning}. For more
details on the implementation and training of the layer functions, see
\cite{Bowman:Potts:Manning:2014}.

@@

This model used here differs from the one described in that work in
two ways. First, because the inputs are single terms, we do not use
a composition function here. Second, for our experiment on WordNet data, 
we introduce an additional neural network layer between the embedding 
input and the comparison function, which facilitates initializing the 
embeddings themselves from pretrained vectors and was found to help 
performance in that setting.

@@

\paragraph{Experiments}
To test our models' ability to learn these inferential patterns, we
create artificial boolean structures in which terms denote sets of entities from
a small domain (e.g., Figure~\ref{lattice-figure}), employ logical
deduction to identify the valid statements, divide those into train
and test sets, and remove from the test set those statements which
cannot be proven from the training statements
(Figure~\ref{unprovable}).
% using the logic depicted in Figure~\ref{lattice-figure}.
%
%
% We test the model's ability to learn this behavior by creating
% artificial data sets of terms which represent sets of numbers. Since
% MacCartney and Manning's set of relations hold between sets as well as
% between sentences, we can use the underlying set structure to generate
% the all of the relations that hold between any pair of these terms, as
% in Figure \ref{lattice-figure}. We train the model defined above on a
% subset of these relations, but rather then presenting the model with a
% pair of tree-structured sentences as inputs, simply present it with
% two single terms, each of which corresponds to a single vector in the
% (randomly initialized) vocabulary matrix $V$, ensuring that the model
% has no information about the terms being compared except the relations
% between them.
%
In our experiments, we create 80 randomly generated sets drawn from a
domain of seven elements. This yields 6400 statements about pairs of
formulae, of which 3200 are chosen as a test set, and that test set is
further reduced to the 2960 examples that can be provably derived from
the training data. We trained both the NN model and the NTN model on
these data
sets.% . In both cases, the models were implemented exactly as
% described in Section~\ref{methods}, but since the items being compared
% are single terms rather than full tree structures, the composition
% layer was not involved, and the two models differed only in which
% layer function was used for the comparison layer. We simply present
% the models with two single terms, each of which corresponds to a
% single vector in the (randomly initialized) vocabulary matrix $V$,
% thereby ensuring that the model has no information about the terms
% being compared except the relations between them.

@@

These results are fairly straightforward to interpret. The NTN model
was able to accurately encode the relations between the terms in the
geometric relations between their vectors, and was able to then use
that information to recover relations that were not overtly included
in the training data. In contrast, the NN was able to achieve this
behavior only incompletely. It is possible but not likely that it
could be made to find a good solution with further optimization on
different learning algorithms, or that it would do better on a larger
universe of sets, for which there would be a larger set of training
data to learn from, but the NTN is readily able to achieve these
effects in the setting discussed here. The fact that both models performed
well above chance on the unprovable examples indicates that both were able
to identify statistical cues to the properties of the underlying boolean structure,
but that information of this kind is not sufficient to yield the near-perfect
accuracy that we saw on the provable data.

@@

We extract three types of relation from WordNet. \ii{Hypernym} and
\ii{hyponym} can be represented directly in the WordNet graph
structure, and correspond closely to the $\natrev$ and $\natfor$
relations from natural logic. As in natural logic, these relations are
mirror images of one another: if \ii{dog} is a hyponym of \ii{animal}
(perhaps indirectly by way of \ii{canid}, \ii{mammal}, etc.), then
\ii{animal} is a hypernym of \ii{dog}. We also extract \ii{coordinate}
terms, which share a direct hypernym, like \ii{dalmatian}, \ii{pug},
and \ii{puppy}, which are all direct hyponyms of \ii{dog}.  Coordinate
terms tend to exclude one another, thereby providing a loose approximation 
of the natural logic exclusion relation $\natalt$. WordNet antonyms are both too narrowly defined
and too rare to use for this purpose. WordNet defines its relations over
sets of synonyms, rather than over individual terms, so we do not
include a \ii{synonym} or \ii{equivalent} relation, but rather
consider only one member of each set of synonyms. Word pairs which do
not fall into these three relations are not included in the data set.

>>>>

Natural logic aligns well with distributed (e.g., vector)
representations, which also naturally model meaning relationally.
Distributed representations have been used successfully in a wide
array of sophisticated language tasks, including part of speach tagging,
\cite{collobert2011natural}, semantic role labeling \cite{collobert2011natural},
sentiment analysis \cite{socher2011semi}, and translation \cite{sutskever2014sequence} 
among many others.
However, it remains an open question whether it is possible to train
such representations to support the rich, diverse logical reasoning
captured by natural logic; while they excel at synonymy (similarity),
the results are more mixed for entailment, contradiction, and mutual
consistency.  Using the natural logic of \cite{maccartney2009extended}
as our formal model, we address this open question using two neural
network-based models for learning embeddings: plain neural networks
and neural tensor networks (NTNs).  The natural logic is built from
the seven relations defined in Table~\ref{b-table}. Its formal
properties are now well-understood \cite{Icard:Moss:2013:LILT}, so it
provides a rigorous set of goals for our neural models. To keep the
discussion manageable, we limit attention to experiments involving the
lexicon; for a more extended treatment of complex expressions
involving logical connectives and quantifiers, see
\cite{Bowman:Potts:Manning:2014}. % CITET

@@

\begin{table*}[tp]
  \centering\small
  \setlength{\tabcolsep}{15pt}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabular}{l c l l} 
    \toprule
    Name & Symbol & Set-theoretic definition & Example \\ 
    \midrule
    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle $\natfor$ reptile}  \\ 
    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile $\natrev$ turtle}  \\ 
    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch $\nateq$ sofa} \\ 
    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle $\natalt$ warthog} \\ 
    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able $\natneg$ unable} \\
    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal $\natcov$ non-turtle} \\ 
    independence       & $x \natind y$   & (else) & \ii{turtle $\natind$ pet}\\
    \bottomrule
  \end{tabular}
  \protect\caption{\protect\label{b-table}The seven natural logic relations of \protect\cite{maccartney2009extended}. 
    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
    and the relation $\natind$ applies whenever none of the other six do.} 
\end{table*}

@@

We build embedding-based models using the method of
\cite{Bowman:Potts:Manning:2014}, which is centered on the task of
labeling a pair of words or sentences with one of a small set of
logical relations. Unlike in a classical inference
setting, the model is not given specific premises to reason with at test time, but rather
is expected to encode all of the information that it learns about each term (word) at training
time in the term vectors, and to then use only the information encoded in the relevant pair of
term vectors at test time. 
The architecture that we use, which is limited to
only pairs of single terms, is depicted in
Figure~\ref{sample-figure}. The model represents the two input terms
as embeddings, which are fed into a comparison function based on one
of two types of neural network layer functions to produce a
representation for the relationship between the two terms. This
representation is then fed into a softmax classifier, which outputs a
distribution over possible labels. The entire network, including the
embeddings, is trained using backpropagation with AdaGrad \cite{duchi2011adaptive}
with L2 regularization.

The simpler version of the comparison concatenates the two
input vectors before feeding them into a LReLU \cite{maasrectifier} neural network (NN)
layer.  The more powerful neural tensor network (NTN) version adds an
additional third-order tensor parameter to allow for multiplicative
interactions between the two inputs \cite{chen2013learning}. For more
details on the implementation and training of the layer functions, see
\cite{Bowman:Potts:Manning:2014}.

@@

This model used here differs from the one described in that work in
two ways. First, because the inputs are single terms, we do not use
a composition function here. Second, for our experiment on WordNet data, 
we introduce an additional $\tanh$ neural network layer between the embedding 
input and the comparison function, which facilitates initializing the 
embeddings themselves from pretrained vectors and was found to help 
performance in that setting.

@@

\paragraph{Experiments}
To test our models' ability to learn these inferential patterns, we
create artificial boolean structures in which terms denote sets of entities from
a small domain (e.g., Figure~\ref{lattice-figure}), employ logical
deduction to identify the valid statements, divide those into train
and test sets, and remove from the test set those statements which
cannot be proven from the training statements
(Figure~\ref{unprovable}). 
% using the logic depicted in Figure~\ref{lattice-figure}.
%
%
% We test the model's ability to learn this behavior by creating
% artificial data sets of terms which represent sets of numbers. Since
% MacCartney and Manning's set of relations hold between sets as well as
% between sentences, we can use the underlying set structure to generate
% the all of the relations that hold between any pair of these terms, as
% in Figure \ref{lattice-figure}. We train the model defined above on a
% subset of these relations, but rather then presenting the model with a
% pair of tree-structured sentences as inputs, simply present it with
% two single terms, each of which corresponds to a single vector in the
% (randomly initialized) vocabulary matrix $V$, ensuring that the model
% has no information about the terms being compared except the relations
% between them.
%
In our experiments, we create 80 randomly generated sets drawn from a
domain of seven elements. This yields 6400 statements about pairs of
formulae, of which 3200 are chosen as a test set. For 240 of the test examples, 
there is not sufficient evidence in the training data to support the choice of any 
one relation, and we  exclude these from our primary test set. 
We trained both the NN model and the 
NTN model on these data sets.

@@

These results are fairly straightforward to interpret. The NTN model
was able to accurately encode the relations between the terms in the
geometric relations between their vectors, and was able to then use
that information to recover relations that were not overtly included
in the training data. In contrast, the NN was able to achieve this
behavior only incompletely. It is possible but not likely that it
could be made to find a good solution with further optimization on
different learning algorithms, or that it would do better on a larger
universe of sets, for which there would be a larger set of training
data to learn from, but the NTN is readily able to achieve these
effects in the setting discussed here. The fact that both models performed
well above chance on the unprovable examples indicates that both were able
to identify statistical cues to the properties of the underlying boolean structure,
thereby going beyond what one would expect from a logical proof system. However, 
information of this kind is not sufficient to yield the near-perfect
accuracy that we saw on the provable data.

@@

We extract three types of relation from WordNet. \ii{Hypernym} and
\ii{hyponym} can be represented directly in the WordNet graph
structure, and correspond closely to the $\natrev$ and $\natfor$
relations from natural logic. As in natural logic, these relations are
mirror images of one another: if \ii{dog} is a hyponym of \ii{animal}
(perhaps indirectly by way of \ii{canid}, \ii{mammal}, etc.), then
\ii{animal} is a hypernym of \ii{dog}. We also extract \ii{coordinate}
terms, which share a direct hypernym, like \ii{dalmatian}, \ii{pug},
and \ii{puppy}, which are all direct hyponyms of \ii{dog}.  Coordinate
terms tend to exclude one another, thereby providing a loose approximation 
of the natural logic exclusion relation $\natalt$. WordNet's antonym relation, however, is both too narrowly defined
and too rare to use for this purpose. WordNet defines its relations over
sets of synonyms, rather than over individual terms, so we do not
include a \ii{synonym} or \ii{equivalent} relation, but rather
consider only one member of each set of synonyms. Word pairs which do
not fall into these three relations are not included in the data set.
